Dynamical Properties of Molecular Cloud Complexes in Galaxies at the Epoch of Reionization

### 2019-03-15 WRITEUP/ms.tex
merged w/ pallo comments

\AP{i have to check abstract and Section 5., in particular 5.2; check also what have been moved to the app}

@@

% radiative feedback --> NT pressure
Star formation is modeled using a H$_2$-based prescription of the Schmidt-Kennicutt relation \citep{Krumholz09a}. We adopt stellar tracks from \ncode{starburst99} and include stellar feedback from supernovae and OB/AGB stars to account for energy dissipations. To couple the feedback to the gas, we employ sub-grid modeling for blastwaves, e.g., to account for the potential lost of energy inside the cell during a supernova explosion. The energy dissipated is then injected into the ISM in the form of kinetic and thermal energies. Radiation pressure on the dust and gas is also included (see \citealt{Pallottini17a} for details).

@@

In \Fig{h2density}, we show an example of the volumetric H$_2$ density ($n_{\rm H2}$) distribution of \flower~for a given evolutionary stage which includes contributions within $R\sim$\,3.5\,kpc from the galaxy center.
%
We note that the distribution is almost flat for $n_{\rm H2}\gtrsim1$\,\cc and it samples the range of density where ``clumps/structures'' are found based on the Euler characteristic, i.e., the fourth Minkowsky functional (see \citealt{Pallottini17b}). We identify MCs by applying 10 equally-spaced volumetric H$_2$ density cuts of $\log{(n_{\rm cut}/{\rm cm^{-3})}}\in[-0.5, 1.5]$\eq[0.32, 0.53, 0.88, 1.45, 2.45, 4.08, 6.81, 11.36, 19.00, 31.62]\footnote{We also vary the range of $n_{\rm cut}$ adopted and find no qualitative differences affecting the results and conclusions of this work except in the slope of the CMF (see \Sec{ncut}).} to each snapshot.

@@

% Relating Larson's law to stability and pressure (perhaps move this to other section?).
In general, the virial theorem for a distribution of gas can be written as
\begin{equation}\label{eqn:virial_th_general}
\frac{1}{2}\ddot I = 2(T - T_{\rm ext}) + W,
\end{equation}
where $\ddot I$ is the second time derivative of the Lagrangian moment of inertia, $T$ is the internal energy of the gas, $T_s$ is external pressure support, and $W$ is the gravity energy.
%
Let us specialize to the case of a spherical self-gravitating molecular cloud of mass $M_{\rm cl}$, radius $R$, and velocity dispersion $\sigma$, accounting for both thermal and turbulent velocity dispersion. Neglecting magnetic field and calling $P_{\rm ext}$ the external pressure, \Eq{virial_th_general} can be written as
\begin{equation}
\frac{1}{2}\ddot I = 3 M_{\rm cl} \sigma^2 - 4\pi P_{\rm ext} R - \frac{\Gamma GM_{\rm cl}^2}{R}\,,
\label{eqn:virial}
\end{equation}
where $\Gamma$ is a geometrical factor that is equal to 3/5 for a uniform sphere; in \Eq{virial} the first term on the right hand side (RHS) is the pressure term from internal velocity dispersion, the second term the external pressure term, and the third term is is the gravity term. For the case of simple virial equilibrium (SVE; i.e., equilibrium state without external pressure), \Eq{virial} becomes:
\begin{equation}
\alpha_{\rm vir} \equiv \frac{3\sigma^2}{\Gamma GM_{\rm cl}/R} = \frac{5\sigma^2R}{GM_{\rm cl}}\,,
\label{eqn:alpha}
\end{equation}
which is also the defined as the virial parameter $\alpha_{\rm vir}$.

@@

The onset of gravitational instability is considered to be tightly connected to \SF \citep[e.g.,][]{Kennicutt89a, Wang94a, Li05b, Li06a}. \AP{i think we can partially cut the following} \DL{probably, but I am just being explicit here to help remind myself what am I trying to say.}
From virial theorem, Jeans instability sets the scales at which gravitational potential overcomes the internal energy. Namely, gravitational instabilities dominates at scales $L > L_J$, where $L_J$ is the Jeans length. However, centrifugal force resulting from differential rotation in disk galaxies can inhibit collapse even on these scales. As a result, a disk is thought to be unstable on scales between $L_J < L < L_{\rm rot}$, where $L_{\rm rot}$ is set by the surface density of the disk and the angular velocity of the rotation.

In theoretical and observational works, gravitational instability has been widely discussed in the context of the Toomre stability criterion \citep{Toomre64a, Goldreich65b}, or the Toomre $Q$ parameter (\AP{i would put this parenthesis as a note later on, i.e. close to \Eq{q_eff}} or the proxy, circular velocity to dispersion ratio; $v_{\rm circ}/\sigma$; e.g., \citealt{GarciaBurillo03a, Genzel11a, Kassin12a, Leung19a}).
% https://ned.ipac.caltech.edu/level5/March15/Glazebrook/Glazebrook5.html
%
% from linear analysis
For {\it axisymmetric} mode, the dispersion relation for the growth of density perturbation in a rotating turbulent disk of
finite thickness $h$ is described by
\begin{equation}
\omega^2 = \kappa^2 - \frac{2\pi G \Sigma |k|}{1 + |k| h} + \sigma^2 k^2\,,
\label{eqn:3Ddisp}
\end{equation}
where $k$ is the wavenumber and $\kappa$ is the epicyclic frequency, defined as:
\begin{equation}
\kappa^2\equiv\frac{2\Omega}{R}\frac{d}{dR}\left(R^2\Omega\right)\,,
\label{eqn:kappa}
\end{equation}
\citep{Romeo92a}.
The first term on the RHS of Equation~\ref{eqn:3Ddisp} describes rotation support, whereas the second term describes self-gravity, and the third term describes internal pressure support. The Toomre-Q parameter can be defined as \citep{Toomre64a}:
\begin{equation}
Q_{\rm gas}\equiv\frac{\sigma\kappa}{\pi G \Sigma}\,.
\label{eqn:Q}
\end{equation}
In the thin disk approximation ($kh\ll1$), the solution of \Eq{3Ddisp} is unstable for scales $k$ such that 
$Q < Q_{\rm crit}\simeq1$ (or equivalently $\omega^2 < 0$). Physically, this situation happens when local gravitational instability
overcomes support from pressure and rotation, allowing fragmentation and collapse to happen.
Similarly, for a collisionless fluid --- as an ensemble of stars --- the Toomre $Q$ parameter is defined by \citep{Toomre64a}:
\begin{equation}
Q_{\star} \equiv\frac{\sigma\kappa}{3.36 G \Sigma}\,.
\end{equation}

In our stability analysis, we account for a non-negligible disk thickness and for the combined effect of gas and stars; this is done adopting the approximation for an effective two-component $Q_{\rm eff}$ parameter \citep[i.e.,][see also \citealt{Inoue16a}]{Romeo11a}.
%
The effect of disk thickness modify the $Q$ parameter for gas/star by accounting for the vertical velocity dispersion:
\begin{equation}
T_{x} = \left\{
		\begin{array}{lccr}
			{\displaystyle 0.8 + 0.7\left(\frac{\sigma_{z}}{\sigma_{r}}\right)}      && & \mbox{if\ } \sigma_z \lesssim 0.5 \times \sigma_r \\ [1.25em]
			{\displaystyle 1 + 0.6\left(\frac{\sigma_{z}}{\sigma_{r}}\right)}        & & & \mbox{if\ } \sigma_z \gtrsim 0.5 \times \sigma_r
\\
		\end{array}
	\right.
\end{equation}
and
\begin{equation}
Q^{\rm thick}_{x} = T_{x} Q\,,
\end{equation}
with $x$ indicating either gas or stars.
% Note that w/ thickness, there's the a self-gravity dilution term in the dispersion relation, which in effect reduceds the surface gravity in the vertical direction. As such, it is easier to be stable against gravitaional instability. Thus, the criterion for instability $Q$ is reduced.
%
The combined effect of gas and stars can then be accounted for by writing
\begin{equation}\label{eqn:q_eff}
Q^{-1}_{\rm eff} =  \left\{
				\begin{array}{lccr}
					     {\displaystyle\frac{w}{Q^{\rm thick}_{\rm star}} + \frac{1}{Q^{\rm thick}_{\rm gas}}}      & & & \mbox{if\ }  Q^{\rm thick}_{\rm star} \geq Q^{\rm thick}_{\rm gas} \\ [0.75em]
                                               {\displaystyle\frac{1}{Q^{\rm thick}_{\rm star}} + \frac{w}{Q^{\rm thick}_{\rm gas}}}      & & & \mbox{if\ } Q^{\rm thick}_{\rm star} \leq Q^{\rm thick}_{\rm gas} \\
				\end{array}
			    \right.
\end{equation}
where the relative weight $w$ is defined as:
\begin{equation}
w\equiv\frac{2 \sigma_{\star} \sigma_{\rm gas}}{\sigma_{\star}^2 + \sigma_{\rm gas}^2},
\end{equation}
Conceptually, the thickness of a disk reduces the gravity in the vertical direction, thereby making 
a system easier to obtain stability, and thus, lowering the criterion of the critical Toomre 
$Q_{\rm crit}$ from $\simeq$\,1 to 
0.67 \citep{Goldreich65a}. On the other hand, including the contribution of the stellar 
component promotes gravitational instability, and thus, increasing the criterion of the Toomre $Q_{\rm crit}$.
unless it is dynamically hot to ``compensate''.  
As a rule of thumb, $Q_{\rm crit}\eq1.34$ for $Q_{\rm gas}$\eq$Q_\star$.  % 0.67 * 2
\AP{symbol already used for internal enrgy in \Eq{virial_th_general}; do we already need the virial theorem derivation?}
\DL{this is small $w$. Whether or not we need the Larson's relation stuff, I think those eqns there help understand the motivation
behind relating aa with bb and bb and cc, etc, in the Larson's relations?}

@@

\begin{figure*}
\centering
\includegraphics[trim=30 30 0 0, clip, width=1.1\textwidth]{\figpath/3by2_clumpProp_ss16-27.pdf}
\caption{
Linewidth-size relation (top),
{\bf .... (middle), and blah blah (bottom)}
for MCs (star symbols)
identified in the two most extreme evolutionary stages of \flower\ --- accreting phase (left) and starburst phase (right).
Star symbols are color-coded by the density thresholds ($n_{\rm cut}$). \DL{some missing character after downloading 
to local.. only the case if we saved them as .pdf.. interestingly but not the case if I save as .png}
\label{fig:larsons_single}}
\end{figure*}

@@

For each of the two most extreme evolutionary stages of \flower\ --- accretion and starburst phase (see \Sec{sfh}), we identify a set of MCs.
%
Scaling relations for the MCs identified in the accretion phase are shown in left panels of \Fig{larsons_single} and in the right
panels for the starburst phase.
%
Regardless of the H$_2$ density cuts adopted, the MCs are found to have consistently \AP{which is already said in the last part of previous Sec.; I would put here that part} \DL{?}
high turbulences and surface densities in both stages. These correspond to MCs at the center of the main disk of \flower and occupy the top right corner of the Larson's relation shown in \Fig{larsons_single}.
%
As we increase the density threshold, some of the MCs within the main disk break into multiple sub-MCs. As such, we effectively identify a population of denser molecular structures (see \Fig{MC}).

@@

\begin{figure*}[htbp]
\centering
\includegraphics[trim=0 0 0 0, clip, width=0.8\textwidth]{\figpath/lf16_cloud-mass_size-pc.png}
\caption{
Size-mass relation of MCs identified in the accretion phase of \flower in our simulation (star symbols) compared to observational data of molecular clouds in the Milky Way associated with massive \SF (magenta circles, green stars, blue dots, and black triangles) and empirical relations established based on \obs of the Milky Way. Red line shows the threshold for massive \SF reported by \citet{Kauffmann10b}. Star symbols are color-coded by increasing $n_{\rm cut}$. Literature data are compiled from \citet{Beuther02a, Mueller02a, Hill05a, Motte07a}. The colored lines show the loci expected for various visual extinctions ($A_V$), which corresponds to lines of constant surface density (i.e., Larson's third relation). This representation is motivated by observational studies (see text and e.g., \citealt{Lombardi10a}).
\AP{same color code as in the panels with different ncut}
\label{fig:MR}}
\end{figure*}

@@

The MCs of \flower has dynamics largely similar to those observed in $z$\ssim2 spatially resolved studies of gas-rich star-forming galaxies, in terms of their velocity dispersions, sizes, and gas surface densities (\Fig{larsons_single}; see e.g., \citealt{Swinbank11a}), with sizes of the order of $R\simeq$\,100\,pc and velocity dispersions of $\sigma\simeq$\,20$-$80\,\kms.
%
% velocity dispersion, larsons
%
Their velocity dispersions are also comparable to those observed in the inner Milky way and nearby gas-rich galaxies (e.g., M64; \citealt{Oka01a, Rosolowsky05a, Heyer09a}), which lies along the locus of $\sigma\propto R^{0.56}$. Such high velocity dispersions and surface densities are expected since they are located in the nuclear regions of the galaxy, where the potential well is also deep. The higher velocity dispersions
(see footnote~\ref{ftn:veldisp} on page \pageref{ftn:veldisp}) of these MCs can also be understood as they have experienced more recent episodes of \SF as \flower is assembling its stellar mass (these MCs also have higher stellar-to-gas mass ratios of $\sim$60).
%
In fact, by $z\simeq$\,7.2 (\AP{i would always refer to the evolutionary stage name, once in defined in Fig. 1; we might want to use italic to underline it is a definition} i.e., accretion phase; see \Fig{SFH}a), \flower has assembled a stellar mass of $M_*$\eq7.5\E{9}\,\Msun. Thus, the higher velocity dispersion is due in part to the stronger stellar feedback compared to e.g., MCs in its satellite galaxies ($\sigma\approx$10\,\kms).

The high pressure observed in \flower is comparable to what has been observed in local ultra-luminous IR galaxies (ULIRGs); however, the molecular clouds in the latter are concentrated within their central regions and have typical sizes of $\sim$70-100\,pc and masses on the order of $\sim$10$^9$\,\Msun \citep{Downes98a, Sakamoto08a}. In our simulated galaxy, the high pressure MCs of \flower are found throughout the disk. This difference likely stems from the different physical mechanisms giving rise to the formation (and thus the nature) of these molecular structures.
%
For instance, in local ULIRGs, they are likely form by shock compression and cloud-cloud collision after large amount of gas from the pair of progenitor galaxies are being funneled toward the central region \citep{Tan00a, Wu18a}.
% Tasker09a
On the other hand, the highly turbulent structures of \flower likely result from constant cold gas accretion onto the entire galaxy \AP{i do not fully grasp the following point} \DL{doesn't it take a long time for gas to lose angular 
momentum to fall onto the central region of a galaxy? Given turbulent it is across the entire galaxy \Fig{sigma}, 
isn't more conceivable that cold gas accretion is the dominant reason to the turbulent structures in \flower?} 
(without requiring gas to be funneled into the central region).
This latter mechanism may also be the dominant mode for forming the highly supersonic massive MCs observed in gas-rich star-forming galaxies at $z$\ssim2 (see also e.g., \citealt{Swinbank11a}) --- but the observed size scales could also result from resolution limit where the true physical scales of MCs in $z$\,\ssim2 galaxies could be smaller in reality.
%
The constant inter-cloud collision as gas is being accreted onto the main galaxy would also render
the MCs to be gravitationally unbound (see e.g., \citealt{Dobbs11a}), which is
consistent with the high $\alpha_{\rm vir}$ found for the MCs in the main disk of \flower.

@@

% alpha parameters
As shown in \Fig{larsons_single},
the virial parameter of the sub-MCs is about $\alpha_{\rm vir}\gtrsim10$, which is considerably lower than their ``parent'' MCs found in the main disk of \flower ($\alpha_{\rm vir}\gtrsim100$).
%
For the MCs in the satellite galaxies, a lower virial parameter is found compared to the (sub-)MCs in the main disk of \flower, with $\alpha_{\rm vir}\approx$\,2.
Differences in their $\alpha_{\rm vir}$ most likely result from the weaker stellar feedback in the satellite galaxies as they have experienced less episodes of \SF compared to \flower (i.e., lower stellar-to-gas mass ratios of $\sim$0.1).

@@

For MCs in the satellite galaxies, we find even lower pressures and virial parameters, which are comparable to those found in clouds formed in the galactic disks  ($\alpha_{\rm vir}$\eq0.2$-$10; \citealt{Dobbs08a, Tasker09a}).
This suggests that MCs identified in the satellite galaxies are likely to be collapsing structures, and thus,\AP{i suspect a low or null \SF in the satellites, i.e. some of them should have $f_gas = 1$, i.e. no stars} \SF is expected to continue as gas from satellite galaxies is being accreted onto the main galaxies during the EoR.

 \DL{should probably move to somwehere later in this sec}
Maps of the Toomre $Q$ parameters for the gas, stars, and the effective two-component $Q_{\rm eff}$ are shown in \Fig{Qeff}
(see also \Fig{h2density} for example of MCs identified in this evolutionary stage of \flower
using different $n_{\rm cut}$). Fragmentation are expected in regions of low Toomre $Q$ values.
Close resemblance of the $Q_{\star}$ and $Q_{\rm eff}$ maps (especially in the central region of \flower)
indicates that contributions from the stellar component play an important role in
governing the stability of the MCs against $m$\eq0 perturbations.
This can be understood since stars in \flower dominates the central part of the galaxy in mass,
so their gravitational potential provides a non-negligible contribution to the instability.
Similarly, contribution from the thickness of the disk is important in \flower since its disk is not thin (with a
scale height to radius ratio of $h/r_{\rm gal}$\ssim150\,pc/1\,kpc\,$\simeq$\,0.15) and since \flower is more of a warped disk. That is, some MCs are found in regions of $Q_{\rm gas}$\ssim1, 
which would suggest that they are gravitationally unstable;
however, when including the effects due to its stellar potential and the
thickness of its disk (via $\sigma_z$), some of these MCs are consistent with $Q_{\rm eff} > 1$.
This highlights the importance of accounting for stellar contribution and disk thickness 
when examining the stability of
molecular gas structures, especially in the relatively evolved and enriched systems at high redshifts
that are preferentially being imaged at high resolution with ALMA. 
One should be caution when interpreting results based on gas-only Toomre $Q$ analysis.
In the outer regions of \flower, on the other hand, $Q_{\rm eff}$ resembles $Q_{\rm gas}$, with 
both $\log{Q_{\rm gas}}$ and $\log{Q_{\rm eff}} < -1$. 
Notably, these are the more gas-rich regions (see \Fig{sigma}).

@@

We study the dynamical properties of molecular clouds complexes and their temporal evolution in a \z$\sim$\,6 prototypical galaxy
at the EoR using state-of-the-art cosmological zoom-in simulation (\ncode{Serra}),
which includes a chemical network to determine the formation of molecular
hydrogen, heating and cooling of the ISM by metals, and stellar feedback.
We use a clump-finding algorithm and a set of H$_2$ volumetric densities
to identify MCs and their sub-structures in the main galaxy of the
simulated 20\,Mpc~$h$\pmOne box --- \flower\ --- and in its satellites.
We decompose the molecular structures into non-overlapping tiles
by identifying a set of different density contours at different evolutionary stages.
Using volumetric H$_2$ is essentially similar to identifying molecular structures using
different contour levels/surfaces on surface brightness maps/cubes of molecular line tracers (e.g., CO, CS, NH$_3$),
since the line luminosity scales with the molecular gas density (modulo optical depth effects).
We extract properties such as mass, size, Mach number, velocity dispersion, gas surface density, and virial parameter for each MC and
compare them with those observed in the Milky Way disk, the Galactic center,
and gas-rich starburst galaxies in the local Universe and at the peak epoch of cosmic \SF.
We also examine their properties at the different evolutionary stages of \flower and compare
them with observations.

>>>>

@@

% radiative feedback --> NT pressure
Star formation is modeled using a H$_2$-based prescription of the Schmidt-Kennicutt relation \citep{Krumholz09a}. We adopt stellar tracks from \ncode{starburst99} and include stellar feedback from supernovae (SN) and OB/AGB stars to account for energy dissipations. To couple the feedback to the gas, we employ sub-grid modeling for blastwaves, e.g., to account for the potential lost of energy inside the cell during a supernova explosion. The energy dissipated is then injected into the ISM in the form of kinetic and thermal energies. Radiation pressure on the dust and gas is also included (see \citealt{Pallottini17a} for details).

@@

In \Fig{h2density}, we show an example of the volumetric H$_2$ density ($n_{\rm H2}$) distribution of \flower~for a given evolutionary stage which includes contributions within $R\sim$\,3.5\,kpc from the galaxy center.
%
We note that the distribution is almost flat for $n_{\rm H2}\gtrsim1$\,\cc and it samples the range of density where ``clumps/structures'' are found based on the Euler characteristic, i.e., the fourth Minkowsky functional (see \citealt{Pallottini17b}). We identify MCs by applying 10 equally-spaced volumetric H$_2$ density cuts of $\log{(n_{\rm cut}/{\rm cm^{-3})}}\in[-0.5, 1.5]$\eq[0.32, 0.53, 0.88, 1.45, 2.45, 4.08, 6.81, 11.36, 18.96, 31.62]\footnote{We also vary the range of $n_{\rm cut}$ adopted and find no qualitative differences affecting the results and conclusions of this work (see \Sec{ncut}).} to each snapshot.

@@

% Relating Larson's law to stability and pressure (perhaps move this to other section?).
In general, the virial theorem for a distribution of gas can be written as
\begin{equation}\label{eqn:virial_th_general}
\frac{1}{2}\ddot{\mathcal{I}} = 2(\mathcal{T} - \mathcal{T}_{\rm ext}) + \mathcal{W},
\end{equation}
where $\ddot{\mathcal{I}}$ is the second time derivative of the Lagrangian moment of inertia, 
$\mathcal{T}$ is the internal energy of the gas, 
$\mathcal{T}_{\rm ext}$ is external pressure support, and $\mathcal{W}$ is the gravitational energy.
%
Let us specialize to the case of a spherical self-gravitating molecular cloud of mass $M_{\rm cl}$, radius $R$, and velocity dispersion $\sigma$, accounting for both thermal and turbulent velocity dispersion. Neglecting magnetic field and calling $P_{\rm ext}$ the external pressure, \Eq{virial_th_general} can be written as
\begin{equation}
\frac{1}{2}{\mathcal{I}} = 3 M_{\rm cl} \sigma^2 - 4\pi P_{\rm ext} R - \frac{\Gamma GM_{\rm cl}^2}{R}\,,
\label{eqn:virial}
\end{equation}
where $\Gamma$ is a geometrical factor that is equal to 3/5 for a uniform sphere; in \Eq{virial} the first term on the right hand side (RHS) is the pressure term from internal velocity dispersion, the second term the external pressure term, and the third term is is the gravity term. For the case of simple virial equilibrium (SVE; i.e., equilibrium state without external pressure), \Eq{virial} becomes:
\begin{equation}
\alpha_{\rm vir} \equiv \frac{3\sigma^2}{\Gamma GM_{\rm cl}/R} = \frac{5\sigma^2R}{GM_{\rm cl}}\,,
\label{eqn:alpha}
\end{equation}
which is also the defined as the virial parameter $\alpha_{\rm vir}$.

@@

The onset of gravitational instability is considered to be tightly connected to \SF \citep[e.g.,][]{Kennicutt89a, Wang94a, Li05b, Li06a}. 
From virial theorem, Jeans instability sets the scales at which gravitational potential overcomes the internal energy. Namely, gravitational instabilities dominates at scales $L > L_J$, where $L_J$ is the Jeans length. However, centrifugal force resulting from differential rotation in disk galaxies can inhibit collapse even on these scales. As a result, a disk is thought to be unstable on scales between $L_J < L < L_{\rm rot}$, where $L_{\rm rot}$ is set by the surface density of the disk and the angular velocity of the rotation.

In theoretical and observational works, gravitational instability has been widely discussed in the context of the Toomre $Q$ stability criterion \citep{Toomre64a, Goldreich65b} or its observable proxy 
$v_{\rm circ}/\sigma$, i.e., the circular velocity to dispersion ratio \citep[e.g.,][]{GarciaBurillo03a, Genzel11a, Kassin12a, Leung19a}.
% https://ned.ipac.caltech.edu/level5/March15/Glazebrook/Glazebrook5.html
%
% from linear analysis
For {\it axisymmetric} mode, the dispersion relation for the growth of density perturbation in a rotating turbulent disk of
finite thickness $h$ is described by
\begin{equation}
\omega^2 = \kappa^2 - \frac{2\pi G \Sigma |k|}{1 + |k| h} + \sigma^2 k^2\,,
\label{eqn:3Ddisp}
\end{equation}
where $k$ is the wavenumber and $\kappa$ is the epicyclic frequency, defined as:
\begin{equation}
\kappa^2\equiv\frac{2\Omega}{R}\frac{d}{dR}\left(R^2\Omega\right)\,,
\label{eqn:kappa}
\end{equation}
\citep{Romeo92a}.
The first term on the RHS of Equation~\ref{eqn:3Ddisp} describes rotation support, whereas the second term describes self-gravity, and the third term describes internal pressure support. The Toomre $Q$ parameter for the gas component can be defined as \citep{Toomre64a}:
\begin{equation}
Q_{\rm gas}\equiv\frac{\sigma_{\rm gas}\kappa}{\pi G \Sigma_{\rm gas}}\,.
\label{eqn:Q}
\end{equation}
In the thin disk approximation ($kh\ll1$), the solution of \Eq{3Ddisp} is unstable for scales $k$ such that 
$Q < Q_{\rm crit}\simeq1$ (or equivalently $\omega^2 < 0$). Physically, this situation happens when local gravitational instability
overcomes support from pressure and rotation, allowing fragmentation and collapse to happen.
Similarly, for a collisionless fluid --- as an ensemble of stars --- the Toomre $Q$ parameter is defined by \citep{Toomre64a}:
\begin{equation}
Q_{\star} \equiv\frac{\sigma_{\star}\kappa}{3.36 G \Sigma_{\star}}\,.
\end{equation}

In our stability analysis, we account for a non-negligible disk thickness and for the combined effect of gas and stars; this is done adopting the approximation for an effective two-component $Q_{\rm eff}$ parameter \citep[i.e.,][see also \citealt{Inoue16a}]{Romeo11a}.
%
The effect of disk thickness modify the $Q$ parameter for gas/star by accounting for the vertical velocity dispersion:
\begin{equation}
T_{x} = \left\{
		\begin{array}{lccr}
			{\displaystyle 0.8 + 0.7\left(\frac{\sigma_{z}}{\sigma_{r}}\right)}      && & \mbox{if\ } \sigma_z \lesssim 0.5 \times \sigma_r \\ [1.25em]
			{\displaystyle 1 + 0.6\left(\frac{\sigma_{z}}{\sigma_{r}}\right)}        & & & \mbox{if\ } \sigma_z \gtrsim 0.5 \times \sigma_r
\\
		\end{array}
	\right.
\end{equation}
and
\begin{equation}
Q^{\rm thick}_{x} = T_{x} Q\,,
\end{equation}
with $x$ indicating either gas or stars.
% Note that w/ thickness, there's the a self-gravity dilution term in the dispersion relation, which in effect reduceds the surface gravity in the vertical direction. As such, it is easier to be stable against gravitaional instability. Thus, the criterion for instability $Q$ is reduced.
%
The combined effect of gas and stars can then be accounted for by writing
\begin{equation}\label{eqn:q_eff}
Q^{-1}_{\rm eff} =  \left\{
				\begin{array}{lccr}
					     {\displaystyle\frac{w}{Q^{\rm thick}_{\star}} + \frac{1}{Q^{\rm thick}_{\rm gas}}}      & & & \mbox{if\ }  Q^{\rm thick}_{\star} \geq Q^{\rm thick}_{\rm gas} \\ [0.75em]
                                               {\displaystyle\frac{1}{Q^{\rm thick}_{\star}} + \frac{w}{Q^{\rm thick}_{\rm gas}}}      & & & \mbox{if\ } Q^{\rm thick}_{\star} \leq Q^{\rm thick}_{\rm gas} \\
				\end{array}
			    \right.
\end{equation}
where the relative weight $w$ is defined as:
\begin{equation}
w\equiv\frac{2 \sigma_{\star} \sigma_{\rm gas}}{\sigma_{\star}^2 + \sigma_{\rm gas}^2},
\end{equation}
Conceptually, the thickness of a disk reduces the gravity in the vertical direction, thereby making 
a system easier to obtain stability, and thus, lowering the criterion of the critical Toomre 
$Q_{\rm crit}$ from $\simeq$\,1 to 
0.67 \citep{Goldreich65a}. On the other hand, including the contribution of the stellar 
component promotes gravitational instability, and thus, increasing the criterion of the Toomre $Q_{\rm crit}$.
unless it is dynamically hot to ``compensate''.  
As a rule of thumb, $Q_{\rm crit}\eq1.34$ for $Q_{\rm gas}$\eq$Q_\star$.  % 0.67 * 2

@@

\begin{figure*}
\centering
\includegraphics[trim=30 30 0 0, clip, width=1.1\textwidth]{\figpath/3by2_clumpProp_ss16-27.pdf}
\caption{
Linewidth-size relation (top),
{\bf .... (middle), and blah blah (bottom)}
for MCs (star symbols)
identified in the two most extreme evolutionary stages of \flower\ --- accreting phase (left) and starburst phase (right).
Star symbols are color-coded by the density thresholds ($n_{\rm cut}$). 
\label{fig:larsons_single}}
\end{figure*}

@@

For each of the two most extreme evolutionary stages of \flower\ --- accretion and starburst phase (see \Sec{sfh}), we identify a set of MCs.
%
Scaling relations for the MCs identified in the accretion phase are shown in left panels of \Fig{larsons_single} and in the right
panels for the starburst phase.
%
Regardless of the H$_2$ density cuts adopted, the MCs are found to have consistently 
high turbulences and surface densities in both stages. These correspond to MCs at the center of the main disk of \flower and occupy the top right corner of the Larson's relation shown in \Fig{larsons_single}.
As we increase the density threshold, some of the MCs within the main disk break into multiple sub-MCs. As such, we effectively identify a population of denser molecular structures (see \Fig{MC}).

@@

\begin{figure*}[htbp]
\centering
\includegraphics[trim=0 0 0 0, clip, width=0.8\textwidth]{\figpath/lf16_cloud-mass_size-pc.png}
\caption{
Size-mass relation of MCs identified in the accretion phase of \flower in our simulation (star symbols) compared to observational data of molecular clouds in the Milky Way associated with massive \SF (magenta circles, green stars, blue dots, and black triangles) and empirical relations established based on \obs of the Milky Way. Red line shows the threshold for massive \SF reported by \citet{Kauffmann10b}. Star symbols are color-coded by increasing $n_{\rm cut}$. Literature data are compiled from \citet{Beuther02a, Mueller02a, Hill05a, Motte07a}. The colored lines show the loci expected for various visual extinctions ($A_V$), which corresponds to lines of constant surface density (i.e., Larson's third relation). This representation is motivated by observational studies (see text and e.g., \citealt{Lombardi10a}).
The star symbols are color-coded by $n_{\rm cut}$, same as \Fig{larsons_single}.
\label{fig:MR}}
\end{figure*}

@@

The MCs of \flower has dynamics largely similar to those observed in $z$\ssim2 spatially resolved studies of gas-rich star-forming galaxies, in terms of their velocity dispersions, sizes, and gas surface densities (\Fig{larsons_single}; see e.g., \citealt{Swinbank11a}), with sizes of the order of $R\simeq$\,100\,pc and velocity dispersions of $\sigma\simeq$\,20$-$80\,\kms.
%
% velocity dispersion, larsons
%
Their velocity dispersions are also comparable to those observed in the inner Milky way and nearby gas-rich galaxies (e.g., M64; \citealt{Oka01a, Rosolowsky05a, Heyer09a}), which lies along the locus of $\sigma\propto R^{0.56}$. Such high velocity dispersions and surface densities are expected since they are located in the nuclear regions of the galaxy, where the potential well is also deep. The higher velocity dispersions
(see footnote~\ref{ftn:veldisp} on page \pageref{ftn:veldisp}) of these MCs can also be understood as they have experienced more recent episodes of \SF as \flower is assembling its stellar mass (these MCs also have higher stellar-to-gas mass ratios of $\sim$60).
%
In fact, by the accretion stage (see \Fig{SFH}a) at $z\simeq$\,7.2, 
\flower has assembled a stellar mass of $M_*$\eq7.5\E{9}\,\Msun. Thus, the higher velocity dispersion is due in part to the stronger stellar feedback compared to e.g., MCs in its satellites ($\sigma\approx$10\,\kms).

The high pressure observed in \flower is comparable to what has been observed in local ultra-luminous IR galaxies (ULIRGs); however, the molecular clouds in the latter are concentrated within their central regions and have typical sizes of $\sim$70-100\,pc and masses on the order of $\sim$10$^9$\,\Msun \citep{Downes98a, Sakamoto08a}. In our simulated galaxy, the high pressure MCs of \flower are found throughout the disk. This difference likely stems from the different physical mechanisms giving rise to the formation (and thus the nature) of these molecular structures.
%
For instance, in local ULIRGs, they are likely form by shock compression and cloud-cloud collision after large amount of gas from the pair of progenitor galaxies are being funneled toward the central region \citep{Tan00a, Wu18a}.
% Tasker09a
On the other hand, the highly turbulent structures of \flower likely result from 
extra-planar flows that are due to gas accretion onto the galaxy and SN-driven outflows \citep{gallerani:2018,kohandel:2019}.
The presence of such extra-planar flows may also be the 
dominant mode for forming the highly supersonic massive MCs observed in gas-rich star-forming galaxies at $z$\ssim2 (see also e.g., \citealt{Swinbank11a}) --- but the observed size scales could also result from resolution limit where the true physical scales of MCs in $z$\,\ssim2 galaxies could be smaller in reality.
%
The constant inter-cloud collision as gas is being accreted onto the main galaxy would also render
the MCs to be gravitationally unbound (see e.g., \citealt{Dobbs11a}), which is
consistent with the high $\alpha_{\rm vir}$ found for the MCs in the main disk of \flower.

@@

% alpha parameters
As shown in \Fig{larsons_single},
the virial parameter of the sub-MCs is about $\alpha_{\rm vir}\gtrsim10$, which is considerably lower than their ``parent'' MCs found in the main disk of \flower ($\alpha_{\rm vir}\gtrsim100$).
%
For the MCs in the satellite galaxies, a lower virial parameter is found compared to the (sub-)MCs in the main disk of \flower, with $\alpha_{\rm vir}\approx$\,2.
Differences in their $\alpha_{\rm vir}$ most likely result from the weaker stellar feedback in the satellites as they have experienced less episodes of \SF compared to \flower (i.e., lower stellar-to-gas mass ratios of $\sim$0.1).

@@

For MCs in the satellite galaxies, we find even lower pressures and virial parameters, which are comparable to those found in clouds formed in the galactic disks  ($\alpha_{\rm vir}$\eq0.2$-$10; \citealt{Dobbs08a, Tasker09a}).
This suggests that MCs identified in the satellite galaxies are likely to be collapsing structures, and thus, \SF is expected to continue as gas from satellite galaxies is being accreted onto the main galaxies during the EoR.

\subsection{Stability of MCs}

@@

We study the dynamical properties of molecular clouds complexes and their temporal evolution in a \z$\sim$\,6 prototypical galaxy, \flower,
at the EoR using state-of-the-art cosmological zoom-in simulation (\ncode{Serra}),
which includes a chemical network to determine the formation of molecular
hydrogen, heating and cooling of the ISM by metals, and detailed stellar feedback.
We use a clump-finding algorithm and a set of H$_2$ volumetric densities
to identify MCs and their sub-structures in \flower and its satellites.
We decompose the molecular structures into non-overlapping tiles
by identifying a set of different density contours at different evolutionary stages.
Such identification method is similar to MC identification based on 
contour levels/surfaces of surface brightness maps/cubes of molecular line tracers (e.g., CO, CS, NH$_3$),
since the line luminosity scales with the molecular gas density (modulo optical depth effects).

We extract properties such as mass, size, Mach number, velocity dispersion, gas surface density, and virial parameter for each MC and we perform a Toomre-$Q$ stability analysis. The properties of the MCs in \flower are compared with observed MCs in the Milky Way disk, the Galactic center, and gas-rich starburst galaxies in the local Universe and at the peak epoch of cosmic \SF. This is done for different evolutionary stages of \flower.
